### 2016-05-02 ch03-lrm/nonparametric.tex
Most comments of Johan on CH03

Here, we focus on using local modeling to separate the different terms, i.e.
\begin{itemize}
  \item the transfer function $\true{G}$,
  \item the leakage term $T$, and
  \item the noise term $V$.
\end{itemize}
The main term of interest, however, is the transfer function $\true{G}$ and a secondary concern is estimating the noise variance $\sigma_{V}$.

@@

The \gls{LPM} can be understood as a specific case of the \gls{LRM}: for  $\order{A} = 0$, the \gls{LRM} reduces to the \gls{LPM}.
As such, we use the notation $\lpm{\numel{W}, \order{B}, \order{C}}$ as a synonym for $\lrm{\nWind, \order{B}, 0, \order{C}}$.

@@

\begin{remark}
Computing the inverse of this Hessian cannot, to the knowledge of the author, be reduced to a form that yields more insight.
However, if the non-diagonal blocks can be neglected, this would lead to an inverse that is also block-diagonal:
\begin{equation}
   \begin{bmatrix}
    \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_A \pd \theta_A } &
    \deemph{0} & 
    \deemph{0} \\
    \deemph{0} &
    \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_B \pd \theta_B } & 
    \deemph{0} \\
    \deemph{0} &
    \deemph{0} & 
    \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_C \pd \theta_C }
  \end{bmatrix}^{-1}
  \!\!\!=
  \begin{bmatrix}
    \left( \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_A \pd \theta_A }\right)^{-1} &
    \deemph{0} & 
    \deemph{0} \\
    \deemph{0} &
    \left( \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_B \pd \theta_B }\right)^{-1} & 
    \deemph{0} \\
    \deemph{0} &
    \deemph{0} & 
    \left( \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_C \pd \theta_C }\right)^{-1}
  \end{bmatrix}
  % \text{.}
\end{equation}
\end{remark}
\paragraph{Bias on the \glsentrydesc{FRF}}
As a first order approximation, the bias on the \gls{FRF} can be written as
\begin{equation}
  \Delta \LocalModel{G}(\omega_k) 
  =
  \Delta \LocalModel{G}(\dw{0}{k}) 
     \approx 
        \PartialDerivative{\LocalModel{G}(\theta, \dw{0}{k})}
                                          {\theta} 
      \Delta \theta
      \text{.}
\end{equation}
The derivative in that expression can be evaluated easily
\begin{align}
    \PartialDerivative{\LocalModel{G}(\theta, \dW)}{\theta} 
     &=
     \begin{bmatrix}
       \PartialDerivative{\LocalModel{G}(\theta,\dW)}{\theta_A} &
       \PartialDerivative{\LocalModel{G}(\theta,\dW)}{\theta_B} &
       \PartialDerivative{\LocalModel{G}(\theta,\dW)}{\theta_C} 
     \end{bmatrix}
     \\
     \PartialDerivative{\LocalModel{G}(\theta, \dW)}{\theta_A} 
     &=
     - \frac{\LocalModel{B}(\theta, r)}{\LocalModel{A}^2(\theta, r)}
     \begin{bmatrix}
         \dW^1 & \dW^2 & \deemph{\cdots} & \dW^{\order{A}}
     \end{bmatrix}
     \\
     \PartialDerivative{\LocalModel{G}(\theta, \dW)}{\theta_B} 
     &=
     \frac{1}{\LocalModel{A}(\theta, \dW)}
     \begin{bmatrix}
         1 & \dW^1 & \dW^2 & \deemph{\cdots} & \dW^{\order{B}}
     \end{bmatrix}
     \\
     \PartialDerivative{\LocalModel{G}(\theta, \dW)}{\theta_C} 
     &=
     \Zero{1 \times \order{C}+1}
     \text{.}
\end{align}

However, the sparsity of $\Delta\theta$, see \eqref{eq:nparam:lrm:bias:sparse}, can be exploited such that the bias on the \gls{FRF} can be expressed as:
\begin{equation}
  \Delta \LocalModel{G}(\theta,\dW)
  \approx
  \frac{\LocalModel{G}(\theta, \dW)}
           {\LocalModel{A}(\theta, \dW)}
  \begin{bmatrix}
    \dW^1 & \deemph{\cdots} & \dW^{\order{A}}
  \end{bmatrix}
  \Delta \theta_A
  \text{.}
\end{equation}
Equivalently, the relative bias of the \gls{FRF} can be written as
\begin{equation}
\frac{\Delta \LocalModel{G}(\theta,\dW)}{\LocalModel{G}(\theta,\dW)}
\approx
\LocalModel{A}(\theta,\dW)^{-1}
\begin{bmatrix}
    \dW^1 & \deemph{\cdots} & \dW^{\order{A}}
  \end{bmatrix}
  \left(  \PartialDerivative[2]{\ELSCost{\theta}}{ \theta^2 } \right)^{-1}
  \begin{bmatrix}
  \Delta\theta_A\\
  0\\
  0\\
  \end{bmatrix}
  \text{.}
\end{equation}

@@

The \gls{FRF} is estimated using the local modeling approaches indicated in \tabref{tbl:nparam:methods}.

@@

\paragraph{Good signal-to-noise ratios}
The next situation we consider is when $\SNR=60\unit{dB}$ as shown in \figref{fig:nparam:comparison:hiSNR}.
Here, the \gls{LPM} remains severly biased for small datasets ($\nBW < 12$).
However, for sufficiently long experiments ($\nBW > 32$), the bias becomes negligible.
For the \gls{LRM} and \gls{LRIC}, the bias is always much smaller than the variance.
In contrast to \gls{LPM}, the \gls{LPM} already reaches a constant variance level when $\nBW > 10$.
It should also be noted that for the \gls{LRIC}, the variance is almost $10\unit{dB}$ more than the corresponding \gls{LPM} method.
It is hence not advantageous to use the \gls{LRIC} as-is.
For all methods (except the \gls{ETFE} and \gls{LRIC}), the simulations show that the \gls{RMSE} reaches
\begin{equation}
  \lim_{\nBW \to \infty} \RMSE \approx \frac{\SNR}{\Delta \SNR}
\end{equation}
where
\begin{equation}
  \Delta\SNR \isdef \sqrt{\frac{\nth}{\nWind}}
  \label{eq:nparam:deltaSNR}
\end{equation}
accounts for the smoothing action of local modeling (see \tabref{tbl:nparam:methods} for numerical values of $\Delta \SNR$).

@@

\paragraph{Poor signal-to-noise ratios}
In \figref{fig:nparam:comparison:lowSNR} and \figref{fig:nparam:comparison:terribleSNR}, the results are shown for $\SNR=20\unit{dB}$ and $\SNR=10\unit{dB}$ respectively.
Again, it can be seen that \gls{LRIC} is plagued by high variance and is hence not directly usable.
However
At $\SNR=20\unit{dB}$ the \gls{LPM} and the \gls{LRM} is significantly biased for small datasets ($\nBW < 6$), however, the \gls{LRM} still perform better than the \gls{LPM}.
For larger datasets ($\nBW \geq 16$), all methods perform similarly and at a nearly constant level of (variance) error.
The \gls{LRM}, in contrast to the \gls{LPM}, still contains an observable bias in the simulations.
At even lower $\SNR=10\unit{dB}$,  most methods perform quite poorly for short datasets ($\nBW < 6$).
However, here it can be seen that the $\lpm{7,4,4}$ outperforms both the $\lrm{7,1,1,1}$ and $\lrm{7,2,2,2}$ and slightly worse than $\lrm{6,2,2,2}$.
It can be seen that the bias error of the \gls{LPM} drops off more quickly ($\bigO{\nBW^{-3}}$) than the \gls{LRM} bias which drops in approximately $\bigO{\nBW^{-1}}$.

@@

\TODO{introduction}

@@

\paragraph{Plug-in methods}
In plug-in methods, a goodness-of-fit criterion (e.g. the log-likelihood function $\LogLikelihood$ of the estimated local model $\LocalModel{M}$ given the estimation dataset $\LocalVector{Z}$), is penalized with a term accounting for the complexity of the model~\citep{Burnham2002}.
E.g. a common choice is the \gls{AIC}~\citep{Akaike1974} or the \gls{AICc}~\citep{Hurvich1989}, which are defined as
\begin{align}
L_{\mathrm{\AIC}} &=
 - 2  \loglikelihood{\LocalModel{M} | \LocalVector{Z}} 
 + 2 \nth\\
L_{\mathrm{\AICc}} &= 
- 2  \loglikelihood{\LocalModel{M} | \LocalVector{Z}} 
+   \frac{2 \nth\nWind}{\nWind - \nth - 1}
\end{align}
where $\nth$ is the number of  estimated parameters in the model $\LocalModel{M}(\theta)$.
In the context of local modeling, the \gls{AIC} is a poor choice since it relies on asymptotic arguments of the data set, i.e. $\nWind \to \infty$, which is obviously an unwarranted choice in this context.
The best model among a set of $q$ possible models $\Set{M_1, \ldots, M_q}$ is then
\begin{equation}
  M_{\star} \isdef 
  \arg_{M} \min 
  \Set{
    L_{\mathrm{AIC(c)}}(M)
  }
  \text{,}
\end{equation}
which is the only model that is hence retained.

@@

@@

\subsection{LPM with impulse response truncation}
\input{\thisDir/truncation.tex}

\section{Conclusion}
\label{sec:conclusion}
In this chapter, the performance of non-parametric \gls{LPM} and \gls{LRM} estimators were compared to observe the \gls{FRF} of resonant systems.
Since the \gls{LRM} is an inherently biased estimator while the \gls{LPM} is asymptotically unbiased, it is important to compare their behavior.
Hence, expressions for the bias of the \gls{LRM} were derived.
For good \glspl{SNR} in resonant systems, it was seen that the \gls{LRM} is always superior to the \gls{LRM}.
For extremely poor \gls{SNR}, i.e. $\SNR \approx 10\unit{dB}$, the \gls{LPM} may perform slightly better than the \gls{LRM} when short measurement records are available.
In this chapter, the \gls{LRIC} was introduced.
This is an extension of the \gls{LRM} which solves the non-linear optimization problem and should thus be able to avoid a bias altogether.
However, it is seen that this estimator is extremely noisy but seems to provide a smaller bias than the \gls{LRM}.
Since the overall \gls{RMS} error for the \gls{LRIC} is much larger than most other methods, the \gls{LRIC} cannot be used in its current form.

>>>>

Here, we focus on using local modeling to separate the different terms, i.e.
\begin{itemize}
  \item the transfer function $\true{G}$,
  \item the leakage term $T$, and
  \item the noise term $V$.
\end{itemize}
The main term of interest, however, is the transfer function $\true{G}$ and the noise variance $\sigma^2_{V}$.

@@

The \gls{LPM} can be understood as a specific case of the \gls{LRM}: for  $\order{A} = 0$, the \gls{LRM} reduces to the \gls{LPM}.
As such, we use the notation $\lpm{\numel{W}, \order{B}, \order{C}}$ as a synonym for $\lrm{\nWind, \order{B}, 0, \order{C}}$.
In the remainder of this text, we refer to \gls{LRM} as a group of methods which have $\order{A} \neq 0$ to make the distinction with \gls{LPM} more obvious.

@@

\begin{remark}
Computing the inverse of this Hessian cannot, to the knowledge of the author, be reduced to a form that yields more insight.
However, if the non-diagonal blocks can be neglected, this would lead to an inverse that is also block-diagonal:
\begin{equation}
   \begin{bmatrix}
    \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_A \pd \theta_A } &
    \deemph{0} & 
    \deemph{0} \\
    \deemph{0} &
    \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_B \pd \theta_B } & 
    \deemph{0} \\
    \deemph{0} &
    \deemph{0} & 
    \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_C \pd \theta_C }
  \end{bmatrix}^{-1}
  \!\!\!=
  \begin{bmatrix}
    \left( \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_A \pd \theta_A }\right)^{-1} &
    \deemph{0} & 
    \deemph{0} \\
    \deemph{0} &
    \left( \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_B \pd \theta_B }\right)^{-1} & 
    \deemph{0} \\
    \deemph{0} &
    \deemph{0} & 
    \left( \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_C \pd \theta_C }\right)^{-1}
  \end{bmatrix}
  % \text{.}
\end{equation}
\end{remark}
\paragraph{Bias on the \glsentrydesc{FRF}}
As a first order approximation, the bias on the \gls{FRF} can be written as
\begin{equation}
  \Delta \LocalModel{G}(\omega_k) 
  =
  \Delta \LocalModel{G}(\dw{0}{k}) 
     \approx 
        \PartialDerivative{\LocalModel{G}(\dw{0}{k},\theta)}
                                          {\theta} 
      \Delta \theta
      \text{.}
\end{equation}
The derivative in that expression can be evaluated easily
\begin{align}
    \PartialDerivative{\LocalModel{G}(\dW, \theta)}{\theta} 
     &=
     \begin{bmatrix}
       \PartialDerivative{\LocalModel{G}(\dW,\theta)}{\theta_A} &
       \PartialDerivative{\LocalModel{G}(\dW,\theta)}{\theta_B} &
       \PartialDerivative{\LocalModel{G}(\dW,\theta)}{\theta_C} 
     \end{bmatrix}
     \\
     \PartialDerivative{\LocalModel{G}(\dW, \theta)}{\theta_A} 
     &=
     - \frac{\LocalModel{B}(\dW, \theta)}{\LocalModel{A}^2(\dW, \theta)}
     \begin{bmatrix}
         \dW^1 & \dW^2 & \deemph{\cdots} & \dW^{\order{A}}
     \end{bmatrix}
     \\
     \PartialDerivative{\LocalModel{G}(\dW, \theta)}{\theta_B} 
     &=
     \frac{1}{\LocalModel{A}(\dW, \theta)}
     \begin{bmatrix}
         1 & \dW^1 & \dW^2 & \deemph{\cdots} & \dW^{\order{B}}
     \end{bmatrix}
     \\
     \PartialDerivative{\LocalModel{G}(\dW, \theta)}{\theta_C} 
     &=
     \Zero{1 \times \order{C}+1}
     \text{.}
\end{align}

However, the sparsity of $\Delta\theta$, see \eqref{eq:nparam:lrm:bias:sparse}, can be exploited such that the bias on the \gls{FRF} can be expressed as:
\begin{equation}
  \Delta \LocalModel{G}(\dW,\theta)
  \approx
  \frac{\LocalModel{G}(\dW, \theta)}
           {\LocalModel{A}(\dW, \theta)}
  \begin{bmatrix}
    \dW^1 & \deemph{\cdots} & \dW^{\order{A}}
  \end{bmatrix}
  \Delta \theta_A
  \text{.}
\end{equation}
Equivalently, the relative bias of the \gls{FRF} can be written as
\begin{align}
\frac{\Delta \LocalModel{G}(\dw{\omega}{}, \theta)}{\LocalModel{G}(\dw{\omega}{}, \theta)}
&\approx
\LocalModel{A}(\dw{\omega}{}, \theta)^{-1}
\begin{bmatrix}
    \dw{\omega}{}^1 & \deemph{\cdots} & \dw{\omega}{}^{\order{A}}
  \end{bmatrix}
  \evaluate{\left(  \PartialDerivative[2]{\ELSCost{\theta}}{ \theta^2 } \right)^{-1}}{A}
  \PartialDerivative{\ELSCost[n]{\true\theta}}{\theta_A}^{\TT}
  \\
  &=
  \frac{\sigma_V^2}{\nWind}
\begin{bmatrix}
    \dw{\omega}{}^1 & \deemph{\cdots} & \dw{\omega}{}^{\order{A}}
  \end{bmatrix}
  \evaluate{\left(  \PartialDerivative[2]{\ELSCost{\theta}}{ \theta^2 } \right)^{-1}}{A}
  \begin{bmatrix}
     \Sum_{\dW} \dW^1 
     \frac{\LocalModel{A}(\conj{\dW},\conj{\LocalVector{\theta}})}
               {\LocalModel{A}(\dw{\omega}{}, \theta)}\\
  \deemph{\vdots}\\
     \Sum_{\dW} \dW^{\order{A}} 
     \frac{\LocalModel{A}(\conj{\dW},\conj{\LocalVector{\theta}})}
               {\LocalModel{A}(\dw{\omega}{}, \theta)}\\
  \end{bmatrix}
\end{align}
where $\evaluate{\left(  \PartialDerivative[2]{\ELSCost{\theta}}{ \theta^2 } \right)^{-1}}{A}$ denotes the upper left $\order{A}\times\order{A}$ subblock of $\left(  \PartialDerivative[2]{\ELSCost{\theta}}{ \theta^2 } \right)^{-1}$ which pertains to the parameters of $\LocalModel{A}$.
From the leading factor, it can be seen that the bias on the \gls{FRF} is $\bigO{\sigma_V^2}$.
Consequently, one can expect that the bias is inherently linked to the \gls{SNR} of the output signal in the frequency domain.

@@

The \gls{FRF} is estimated using the local modeling approaches indicated in \tabref{tbl:nparam:methods}.
This table also indicate $\Delta\SNR$ which indicates the expected gain in \gls{SNR} if the error is mainly determined by the variance.
Concretely,
\begin{equation}
  \Delta\SNR \isdef \sqrt{\frac{\nth}{\nWind}}
  \label{eq:nparam:deltaSNR}
\end{equation}
accounts for the smoothing action of local modeling.

@@

\paragraph{Good signal-to-noise ratios}
The next situation we consider is when $\SNR=60\unit{dB}$ as shown in \figref{fig:nparam:comparison:hiSNR}.
Here, the \gls{LPM} remains severly biased for small datasets ($\nBW < 12$).
However, for sufficiently long experiments ($\nBW > 32$), the bias becomes negligible.
For the \gls{LRM} and \gls{LRIC}, the bias is always much smaller than the variance.
In contrast to \gls{LPM}, the \gls{LRM} already reaches a constant variance level when $\nBW > 10$.
It should also be noted that for the \gls{LRIC}, the variance is almost $10\unit{dB}$ more than the corresponding \gls{LPM} method.
It is hence not advantageous to use the \gls{LRIC} as-is.
For all methods (except the \gls{ETFE} and \gls{LRIC}), the simulations show that the \gls{RMSE} reaches the noise floor minus $\Delta \SNR$.

@@

\paragraph{Poor signal-to-noise ratios}
In \figref{fig:nparam:comparison:lowSNR} and \figref{fig:nparam:comparison:terribleSNR}, the results are shown for $\SNR=20\unit{dB}$ and $\SNR=10\unit{dB}$ respectively.
Again, it can be seen that \gls{LRIC} is plagued by high variance and is hence not directly usable.
However, its bias for small datasets is much lower than for the other methods.
At $\SNR=20\unit{dB}$ the \gls{LPM} and the \gls{LRM} is significantly biased for small datasets ($\nBW < 6$), however, the \gls{LRM} still perform better than the \gls{LPM}.
For larger datasets ($\nBW \geq 16$), all methods perform similarly and at a nearly constant level of (variance) error.
The \gls{LRM}, in contrast to the \gls{LPM}, still contains an observable bias in the simulations.
At even lower $\SNR=10\unit{dB}$,  most methods perform quite poorly for short datasets ($\nBW < 6$).
However, here it can be seen that the $\lpm{7,4,4}$ outperforms both the $\lrm{7,1,1,1}$ and $\lrm{7,2,2,2}$ and slightly worse than $\lrm{6,2,2,2}$.
It can be seen that the bias error of the \gls{LPM} drops off more quickly ($\bigO{\nBW^{-3}}$) than the \gls{LRM} bias which drops in approximately $\bigO{\nBW^{-1}}$.

@@

\TODO{introduction, probleem, oplossingen die besproken worden}

@@

\paragraph{Plug-in methods}
In plug-in methods, a goodness-of-fit criterion (e.g. the log-likelihood function $\LogLikelihood$ of the estimated local model $\LocalModel{M}$ given the estimation dataset $\LocalVector{Z}$), is penalized with a term accounting for the complexity of the model~\citep{Burnham2002}.
E.g. a common choice is the \gls{AIC}~\citep{Akaike1974} or the \gls{AICc}~\citep{Hurvich1989}, which are defined as
\begin{align}
L_{\mathrm{\AIC}} &=
 - 2  \loglikelihood{\LocalModel{M} | \LocalVector{Z}} 
 + 2 \nth
\label{eq:nparam:AIC}\\
L_{\mathrm{\AICc}} &= 
- 2  \loglikelihood{\LocalModel{M} | \LocalVector{Z}} 
+   \frac{2 \nth\nWind}{\nWind - \nth - 1}
\label{eq:nparam:AICc}
\end{align}
where $\nth$ is the number of  estimated parameters in the model $\LocalModel{M}(\theta)$.
In the context of local modeling, the \gls{AIC} is a poor choice since it relies on asymptotic arguments of the data set, i.e. $\nWind \to \infty$, which is obviously an unwarranted choice in this context.
The best model among a set of $q$ possible models $\Set{M_1, \ldots, M_q}$ is then
\begin{equation}
  M_{\star} \isdef 
  \arg_{M} \min 
  \Set{
    L_{\mathrm{AIC(c)}}(M)
  }
  \text{,}
\end{equation}
which is the only model that is hence retained.

\begin{remark}
From a theoretical point of view, equations \eqref{eq:nparam:AIC} and \eqref{eq:nparam:AICc}, require full knowledge of the likelihood function.
This means specifically that the noise variance $\sigma^{2}_{V}$ needs to be known beforehand.
When the variance is unknown, however, it is more robust to incorporate the \gls{AIC} as a penalty factor on the cost function as indicated in \citet[Section 7.4]{Ljung1999} and \citet[Section  11.3 and 19.7]{Pintelon2012}.
\end{remark}

@@

\subsection{Results}

%TODO: discussie: 1 globale modelorder of verschillende
% voor parametrische schatting: vensterbreedte constant houden, DOF constant houden over alle geselecteerde modellen
% voor niet-parametrisch: orde kan per frequentielijn verschillen, geeft betere smoothing \citep{Thummala2012LPMBW,Fan1996}

@@

\section{LPM with impulse response truncation}
\input{\thisDir/truncation.tex}

\section{Conclusion}
\label{sec:conclusion}
In this chapter, the performance of non-parametric \gls{LPM} and \gls{LRM} estimators were compared to observe the \gls{FRF} of resonant systems.
Since the \gls{LRM} is an inherently biased estimator while the \gls{LPM} is asymptotically unbiased, it is important to compare their behavior.
Hence, expressions for the bias of the \gls{LRM} were derived.
%TODO: tom: "inherently biased": verrgaande uitspraak in relatie met Levy: is normaal ook biased, maar voor ARX niet. Hoe zit dat hier?
For good \glspl{SNR} in resonant systems, it was seen that the \gls{LRM} is always superior to the \gls{LPM}.
For extremely poor \gls{SNR}, i.e. $\SNR \approx 10\unit{dB}$, the \gls{LPM} may perform slightly better than the \gls{LRM} when short measurement records are available.
In this chapter, the \gls{LRIC} was introduced.
This is an extension of the \gls{LRM} which solves the non-linear optimization problem and should thus be able to avoid a bias altogether.
However, it is seen that this estimator is extremely noisy but seems to provide a smaller bias than the \gls{LRM}.
Since the overall \gls{RMS} error for the \gls{LRIC} is much larger than most other methods, the \gls{LRIC} cannot be used in its current form.
